---
title: "EMLM Final Case Study"
author: "Colin Yip"
date: "2024-03-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(caret)
library(ggpubr)
library(lme4)
library(GLMMadaptive)
library(MASS)
```

#### Data loading
```{r}
load("Group_9.RData")
resp_var <- DF$Severity
```


#### Distribution of severity relative to covariates
Checking for distribution of severity/bias in severity between groups
```{r}
with(DF, tapply(Severity, list(Visit),
  mean,
  na.rm = TRUE
))
with(DF, tapply(Severity, list(Visit, sex),
  mean,
  na.rm = TRUE
))
with(DF, tapply(Severity, list(Visit, SNP),
  mean,
  na.rm = TRUE
))
with(DF, tapply(Severity, list(Visit),
  var,
  na.rm = TRUE
))
with(DF, tapply(Severity, list(Visit, sex),
  var,
  na.rm = TRUE
))
with(DF, tapply(Severity, list(Visit, SNP),
  var,
  na.rm = TRUE
))

bar_width <- 0.7

sex_labs <- c("Male", "Female")
names(sex_labs) <- c("male", "female")

sex_sev_dist <- ggplot(data = DF, mapping = aes(
  x = as.factor(Severity),
  fill = as.factor(Severity)
)) +
  scale_fill_manual(values = c("seagreen", "purple")) +
  facet_grid(
    cols = vars(sex),
    labeller = labeller(sex = sex_labs)
  ) +
  geom_bar(
    stat = "count",
    color = "black",
    width = bar_width
  ) +
  ggplot2::theme_bw() +
  labs(
    x = "",
    y = "Count",
    title = "Sex"
  ) +
  theme(
    legend.position = "None",
    plot.title = element_text(hjust = 0.5)
  ) +
  ylim(0, 1300)

period_sev_dist <- ggplot(data = DF, mapping = aes(
  x = as.factor(Severity),
  fill = as.factor(Severity)
)) +
  scale_fill_manual(values = c("seagreen", "purple")) +
  facet_grid(cols = vars(Period)) +
  geom_bar(
    stat = "count",
    color = "black",
    width = bar_width
  ) +
  ggplot2::theme_bw() +
  labs(
    x = "",
    y = "Count",
    title = "Period"
  ) +
  theme(
    legend.position = "None",
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ) +
  ylim(0, 1300)

snp_sev_dist <- ggplot(data = DF, mapping = aes(
  x = as.factor(Severity),
  fill = as.factor(Severity)
)) +
  scale_fill_manual(values = c("seagreen", "purple")) +
  facet_grid(cols = vars(SNP)) +
  geom_bar(
    stat = "count",
    color = "black",
    width = bar_width
  ) +
  ggplot2::theme_bw() +
  labs(
    x = "Severity",
    y = "Count",
    title = "SNP"
  ) +
  theme(
    legend.position = "None",
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ) +
  ylim(0, 1300)

joined_fig <- ggarrange(sex_sev_dist, snp_sev_dist, period_sev_dist,
  ncol = 3
)

annotate_figure(joined_fig,
  top = text_grob("Distribution of Severity Level (0, 1) by Categorical Predictor",
    face = "bold",
    size = 13
  )
)
```

Density of relative Severity factor representation as a function of age bins
```{r}
ggplot(
  data = DF,
  mapping = aes(x = Age, fill = as.factor(Severity))
) +
  facet_grid() +
  scale_fill_manual(values = c("seagreen", "purple")) +
  geom_density(alpha = 0.4) +
  ggplot2::theme_bw() +
  guides(fill = guide_legend(title = "Severity")) +
  labs(
    title = "Age Density by Severity Level",
    x = "Age",
    y = "Density"
  )
```
#### Mean structure of severity over visits
```{r}
# Means over time points
plot_ylim <- ylim(0, 0.55)

sex_grouped_data <- DF |>
  group_by(sex, Visit) |>
  summarize(Severity = mean(Severity))

sex_mean_structure <- ggplot(
  data = sex_grouped_data,
  mapping = aes(
    x = factor(Visit),
    y = Severity,
    group = sex,
    color = sex
  )
) +
  geom_line(stat = "summary", fun = "mean") +
  scale_color_manual(
    values = c("#F8766D", "#00BFC4"),
    labels = c("Male", "Female"),
    guide = guide_legend(direction = "horizontal")
  ) +
  ggplot2::theme_bw() +
  labs(
    title = "Sex",
    x = "",
    y = "Severity"
  ) +
  guides(fill = guide_legend(ncol = 2)) +
  theme(
    legend.position = c(0.6, 1.045),
    legend.margin = margin(c(1, 5, 1, 5)),
    legend.title = element_blank()
  ) +
  plot_ylim

snp_mean_structure <- ggplot(
  data = DF |>
    group_by(SNP, Visit) |>
    summarize(Severity = mean(Severity)),
  mapping = aes(
    x = factor(Visit),
    y = Severity,
    group = factor(SNP),
    color = factor(SNP)
  )
) +
  geom_line(stat = "summary", fun = "mean") +
  scale_color_manual(
    values = c("#f72585", "#7209b7", "#207bff"),
    guide = guide_legend(direction = "horizontal")
  ) +
  ggplot2::theme_bw() +
  labs(
    title = "SNP",
    x = "Visit",
    y = "Severity"
  ) +
  guides(fill = guide_legend(ncol = 3)) +
  theme(
    legend.position = c(0.65, 1.045),
    legend.margin = margin(c(1, 5, 1, 5)),
    legend.title = element_blank(),
    axis.title.y = element_blank()
  ) +
  plot_ylim

period_mean_structure <- ggplot(
  data = DF |>
    group_by(Period, Visit) |>
    summarize(Severity = mean(Severity)),
  mapping = aes(
    x = factor(Visit),
    y = Severity,
    group = Period,
    color = Period
  )
) +
  geom_line(stat = "summary", fun = "mean") +
  scale_color_manual(
    values = c("#e4572e", "#ffc914"),
    guide = guide_legend(direction = "horizontal")
  ) +
  ggplot2::theme_bw() +
  labs(
    title = "Period",
    x = "",
    y = "Severity"
  ) +
  theme(
    legend.position = c(0.75, 1.045),
    legend.margin = margin(c(1, 5, 1, 5)),
    legend.title = element_blank(),
    axis.title.y = element_blank()
  ) +
  plot_ylim

joined_mean_structure_fig <- ggarrange(sex_mean_structure, snp_mean_structure, period_mean_structure,
  ncol = 3
)

annotate_figure(joined_mean_structure_fig,
  top = text_grob("Proportion of Severe = 1 by Visit and Predictor",
    face = "bold",
    size = 13
  )
)
```

### Logistic mixed effects model
```{r}
# Change types of variables to factors
DF$Severity <- as.factor(DF$Severity)
DF$Period <- as.factor(DF$Period)
DF$sex <- as.factor(DF$sex)
DF$SNP <- as.factor(DF$SNP)

# Centering and scaling age, so coefficient interpretation becomes
# what is the change in the response relative to a 1SD change in the predictor
DF$scaled_age <- scale(DF$Age)
```

For `glmer`, Laplace approximation is not usable as it does not behave well with binary data. As such, we will use adaptive Gaussian quadrature where possible instead to approximate the integral $\int f(z)\phi(z)dz\approx\sum_{q=1}^{15}w_qf(z_q), \phi(z)\text{ being the multivariate normal},$ which is the likelihood contribution of each subject.

Approach assumes independent errors.

#### Setting up elaborate mean model and picking a random effects structure
```{r}
# Opting for mixed_model directly, as convergence more likely due to
# aGQ being more flexible to either high dimensional, or broadly more uneven/
# extreme likelihood function surfaces --> glmer ran into high eigenvalue ratios
# for the Hessian for increased interaction terms and/or increasing the number of
# random effects. glmer also ran into gradient errors unless a high
# (~1e5) number of iterations was run --> not very computationally efficient

# Check to see if glmer converges
model.0 <- glmer(
  Severity ~ Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age +
    (1 | id),
  data = DF,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa"),
  nAGQ = 15
)

# A penalized quasi-likelihood model runs into similar issues
model.pql <- glmmPQL(
  Severity ~ Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age,
  random = ~ 1 | id,
  family = binomial,
  data = DF
)

# Skipping checks with Laplace due to computational issues with binary variables

# Linear coefficients, random intercept, maximal
model.1 <- mixed_model(
  Severity ~ Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age,
  random = ~ 1 | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

# Linear coefficients, random intercept, random slope, maximal
model.2 <- mixed_model(
  Severity ~ Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

# Quadratic visit, random intercept only, maximal
model.3 <- mixed_model(
  Severity ~ (Visit + I(Visit^2)) * Period +
    (Visit + I(Visit^2)) * sex +
    (Visit + I(Visit^2)) * SNP +
    (Visit + I(Visit^2)) * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age,
  random = ~ 1 | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

# Quadratic visit, random intercept and random slope, maximal
model.4 <- mixed_model(
  Severity ~ (Visit + I(Visit^2)) * Period +
    (Visit + I(Visit^2)) * sex +
    (Visit + I(Visit^2)) * SNP +
    (Visit + I(Visit^2)) * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

if (all(
  model.1$converged, model.2$converged,
  model.3$converged, model.4$converged
)) {
  print("Achieved solution convergence in all models")
} else {
  print("One of the above models has not converged")
}

# Linear vs. quadratic to determine the correct fixed effects
anova(model.1, model.3)
# H_0: Coefficients for quadratic terms are 0, H_A being that they are != 0
# p-value >= 0.05, so we do not include quadratic terms and select from the linear models

# Linear comp (rand int vs rand int + rand slope)
random_effects_lrt <- anova(model.1, model.2)$LRT
0.5 * pchisq(random_effects_lrt, df = 1, lower.tail = FALSE) +
  0.5 * pchisq(random_effects_lrt, df = 2, lower.tail = FALSE)

# Between the linear models, rand int+rand slope (model.2) is signif. better
# Reject H_0: D=[[\sigma^2_int, 0][0,0]]
# Take H_A: D=[[\sigma^2_int, \sigma_{int,time}][\sigma_{int,time},\sigma^2_time]]

# Move forward with model.2
summary(model.2)
```

#### Simplification of elaborate mean model
```{r}
model.2.coefs <- summary(model.2)$coef_table
signif.coefs <- model.2.coefs[model.2.coefs[, "p-value"] < 0.05, ]
signif.coefs

# We take a look at the interaction terms to evaluate which terms to drop
# Starting with sex:scaled_age
model.2.a <- mixed_model(
  Severity ~ Visit + Period + scaled_age + sex + SNP +
    Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    SNP * scaled_age,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

anova(model.2, model.2.a)
```

Take model.2.a, cannot reject H_0 (H_0 being more restrictive model), as p-value of 0.4495.

Move on to evaluate SNP:scaled_age.
``` {r}

model.2.b <- mixed_model(
  Severity ~ Visit + Period + scaled_age + sex + SNP + 
    Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

anova(model.2.a, model.2.b)
```

Take model.2.b, cannot reject H_0 (H_0 being more restrictive model), as p-value of 0.4347.

Move on to evaluate Period:SNP.
```{r}

model.2.c <- mixed_model(
  Severity ~Visit + Period + scaled_age + sex + SNP + 
    Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * scaled_age +
    sex * SNP,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

anova(model.2.b, model.2.c)
```

Take model 2.b, we reject H_0 (H_0 being more restrictive), as p-value of 0.0343.

Move on to evaluate Visit:SNP.
```{r}
model.2.d <- mixed_model(
  Severity ~ Visit + Period + scaled_age + sex + SNP + 
    Visit * Period +
    Visit * sex +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

anova(model.2.b, model.2.d)
```
Take model 2.d, we cannot reject H_0, although by a thin margin, with p-value of 0.065.

Move on to Visit:Age.
```{r}
model.2.e <- mixed_model(
  Severity ~ Visit + Period + scaled_age + sex + SNP + 
    Visit * Period +
    Visit * sex +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

anova(model.2.d, model.2.e)
```
Take model 2.e, we cannot reject H_0, as p-value of 0.7628.

Move on to Visit:Sex
```{r}
model.2.f <- mixed_model(
  Severity ~ Visit + Period + scaled_age + sex + SNP + 
    Visit * Period +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

anova(model.2.e, model.2.f)
```
Take model 2.f, we cannot reject H_0, as p-value of 0.6905.

Move on to Period:Age.
```{r}
model.2.g <- mixed_model(
  Severity ~ Visit + Period + scaled_age + sex + SNP + 
    Visit * Period +
    Period * sex +
    Period * SNP +
    sex * SNP,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

anova(model.2.f, model.2.g)
```
Take model.2.g, we cannot reject H_0, as p-value of 0.4363.

Move on to Sex:SNP.
```{r}
model.2.h <- mixed_model(
  Severity ~ Visit + Period + scaled_age + sex + SNP + 
    Visit * Period +
    Period * sex +
    Period * SNP,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

anova(model.2.g, model.2.h)
```
Take model.2.h, we cannot reject H_0, as p-value of 0.4846.

Move on to Visit:Period.
```{r}
model.2.i <- mixed_model(
  Severity ~ Visit + Period + scaled_age + sex + SNP + 
    Period * sex +
    Period * SNP,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

anova(model.2.h, model.2.i)
```
Take model.2.i, we cannot reject H_0 as p-value of 0.2898.

This concludes model simplification, as we do not drop SNP, age, sex, period as they are tied to research goals of the study. We now validate that the final model indeed has a more optimal structure via a Likelihood Ratio Test, as the final model is a nested model of the initial `model.2`.
```{r}
# Final model
# Has Visit, Period, scaled_age, Sex, SNP, Period:Sex, and Period:SNP
final.glmm <- model.2.i
# Validation that final.glmm is indeed more performant than elaborate model.2
anova(final.glmm, model.2)
```
The result of the anova is a p-value 0.4323, so we cannot reject H_0. Therefore, the smaller model, `final.glmm`, is indeed more performant. The final model contains the following coefficients: 
* Visit
* Period
* scaled_age
* Sex
* SNP
* Period:Sex
* Period:SNP (specifically, the interaction between Period and the level SNP1 is significant, so we retain the entire interaction term)

```{r}
# TODO: Parameter interpretation for final_model with just signif.coefs
log_or_coefficients <- summary(final.glmm)$coef_table[, "Estimate"]
or_coefficients <- exp(log_or_coefficients)
or_coefficients
```

So the final equation of the logistic mixed effects model is as follows:
$\log(\frac{\pi_{ij}}{1-\pi_{ij}})=(-4.2668+b_{i,0})+(0.6352 + b_{i,1}) * Visit_{i,j}-1.9365*Period_i+0.0306*Age_i-0.0203*ScaledAge_i-0.5689*SNP1_i-1.1015*Period_i*Sex_i+1.5821*Period_i*SNP1_i$ for:
* $i=1,\dots,500$ being the patient index
* $j=0,\dots,7$ being the visit index
* $\pi_{i,j}$ being the probability of severe disease for a patient $i$ and visit $j$

And we assume that the random effects follow a normal distribution as follows:
* $\begin{bmatrix}b_{i,0}\\b_{i,1}\end{bmatrix}\sim N_2(\begin{bmatrix}0\\0\end{bmatrix}, \begin{bmatrix}\sigma^2_{int}&\sigma_{int, Visit}\\\sigma_{int, Visit}&\sigma^2_{Visit}\end{bmatrix})$
With a correlation between the two random effects of $$.

Exponentiating these coefficients, the odds-ratio of each term can be interpreted more directly.

```{r}
# TODO: Interpret CI/se for the coefficients
raw_coeff_ci <- confint(final.glmm)
or_coeff_ci <- exp(raw_coeff_ci)
```

```{r}
# Marginal coefficients and their errors
# TODO: Interpret marginal coeffs/SEs
marginal_coeff <- marginal_coefs(final.glmm, std_errors = T)
raw_marginal_coeff_ci <- confint(marginal_coeff)
or_marginal_coeff_ci <- exp(raw_marginal_coeff_ci)
```

#### Model diagnostics
```{r model_diag_q_stability}
final.glmm4 <- update(final.glmm, nAGQ = 4)
final.glmm5 <- update(final.glmm, nAGQ = 5)
final.glmm10 <- update(final.glmm, nAGQ = 10)
final.glmm20 <- update(final.glmm, nAGQ = 20)

models <- list(
  "nAGQ=4" = final.glmm4, "nAGQ=5" = final.glmm5,
  "nAGQ=10" = final.glmm10, "nAGQ=15" = final.glmm,
  "nAGQ=20" = final.glmm20
)

extract <- function(obj) {
  return(c(fixef(obj), "var_(Intercept)" = obj$D[1, 1], "logLik" = logLik(obj)))
}

nagq_q_stability <- data.frame(sapply(models, extract))
knitr::kable(nagq_q_stability,
  caption = "Parameter and Loglikelihood Stability for Different $Q$ Values",
  col.names=c("Q=4", "Q=5", "Q=10", "Q=15", "Q=20")
)
```

Running `mixed_model` for a value of $Q<4$ on this dataset encounters computational issues as the resulting covariance matrix is not positive definite. As such, only values of $Q>=4$ will be considered for assessing the stability of coefficients and the likelihood of the model.

```{r model_diag_residuals}
snp.labs <- c("SNP=0", "SNP=1", "SNP=2")
names(snp.labs) <- c(0, 1, 2)

period.labs <- c("Period A", "Period B")
names(period.labs) <- c("A", "B")

res_check_df <- DF
fitted_residuals <- residuals(final.glmm, 
                              type="subject_specific")
res_check_df$residuals <- fitted_residuals

ggplot(data=res_check_df,
       mapping=aes(x=Visit,
                   y=residuals)) +
  facet_grid(
    cols = vars(SNP),
    rows = vars(Period),
    labeller = labeller(
      SNP = snp.labs,
      Period = period.labs
  )) +
  geom_point(color = "dodgerblue4")

```


```{r model_diag_random_effect_normality, fig.width=8}
final_random_effects <- data.frame(ranef(final.glmm))

rand_int_plot <- ggplot(
  data = final_random_effects,
  mapping = aes(x = X.Intercept.)
) +
  geom_histogram(aes(y = ..density..),
    color = "black",
    fill = "white",
    lwd = 0.2,
    bins = 40
  ) +
  geom_density(color = "dodgerblue4", ) +
  theme_bw() +
  labs(
    y = "Density",
    x = "Empirical Bayes Estimates",
    title = "EB Estimates of Random Intercepts"
  )

rand_slope_plot <- ggplot(
  data = final_random_effects,
  mapping = aes(x = Visit)
) +
  geom_histogram(aes(y = ..density..),
    color = "black",
    fill = "white",
    lwd = 0.2,
    bins = 40
  ) +
  geom_density(color = "dodgerblue4", ) +
  theme_bw() +
  labs(
    y = "Density",
    x = "Empirical Bayes Estimates",
    title = "EB Estimates of Random Slopes"
  )

ggarrange(rand_int_plot, rand_slope_plot, ncol = 2)
```

Not very normal plots for either, with both skewing slightly to the left. In particular the Random Intercepts have a very long tail to the right, indicating that there are individuals at baseline with a significantly higher probability of severe outcomes to begin with.

```{r model_diag_random_effect_independence, fig.width=5, fig.height=5}
ggplot(
  data = final_random_effects,
  mapping = aes(
    x = X.Intercept.,
    y = Visit
  )
) +
  geom_point(color = "dodgerblue4") +
  labs(
    x = "Random Intercepts",
    y = "Random Slopes",
    title = "Scatterplot of EB Estimates"
  )
```

This shows that there is a potential mild correlation between the two random effects, as can be reflected in the $-0.333$ correlation coefficient in the summary of the final model.

#### Visualizing random and fixed effect structures
```{r glmm_viz_average_conditional_probability_M}
# FIXME: Handle gender diff --> plot as 2 diff lines
data.fitted.mean <- data.frame(
  id = 1,
  Visit = 0:7,
  Period = rep(c(rep("A", 8), rep("B", 8)), 3),
  SNP = factor(0:2),
  scaled_age = mean(DF$scaled_age),
  sex = c(rep("male", 48), rep("female", 48))
)

probs.m <- predict(final.glmm, 
                 newdata = data.fitted.mean, 
                 type_pred = "response", 
                 type = "mean_subject", 
                 re.form = NA, 
                 se.fit = TRUE)
data.fitted.mean$fit <- probs.m$pred
data.fitted.mean$fit_se <- probs.m$se.fit

ggplot(
  data = data.fitted.mean,
  mapping = aes(x = Visit, y = fit, color=sex)
) +
  facet_grid(
    cols = vars(SNP),
    rows = vars(Period),
    labeller = labeller(
      SNP = snp.labs,
      Period = period.labs
    )
  ) +
  geom_line() +
  geom_point() +
  scale_color_manual(values=c("coral2", "dodgerblue4"),
                     labels=c("Female", "Male"),
                     name="Sex") +
  theme_bw() +
  labs(
    x = "Visit",
    y = "Probability of Severe Disease",
    title = "Fitted Probability of Severe Disease for Mean Subject"
  ) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")
```

```{r glmm_viz_subj_specific_probability}
strata_sample <- function(x, n) {
  return(sample(unique(x$id), n))
}

# Randomly sample n subjects from Period A for each SNP level,
# and Period B for each SNP level
n_sample_per_strata <- 10

period_A_data <- DF |> filter(Period == 'A')
A_0_data <- strata_sample(period_A_data |> filter(SNP == 0), n_sample_per_strata)
A_1_data <- strata_sample(period_A_data |> filter(SNP == 1), n_sample_per_strata)
A_2_data <- strata_sample(period_A_data |> filter(SNP == 2), n_sample_per_strata)

period_B_data <- DF |> filter(Period == 'B')
B_0_data <- strata_sample(period_B_data |> filter(SNP == 0), n_sample_per_strata)
B_1_data <- strata_sample(period_B_data |> filter(SNP == 1), n_sample_per_strata)
B_2_data <- strata_sample(period_B_data |> filter(SNP == 2), n_sample_per_strata)

sampled_ids <- c(A_0_data, A_1_data, A_2_data,
                 B_0_data, B_1_data, B_2_data)
data.fitted.ss <- DF |>
  filter(id %in% sampled_ids)
probs <- predict(final.glmm, 
                 newdata = data.fitted.ss, 
                 type_pred = "response", 
                 type = "subject_specific",
                 se.fit = TRUE)
data.fitted.ss$fit <- probs$pred

snp.labs <- c("SNP=0", "SNP=1", "SNP=2")
names(snp.labs) <- c(0, 1, 2)

period.labs <- c("Period A", "Period B")
names(period.labs) <- c("A", "B")

ggplot(
  data = data.fitted.ss,
  mapping = aes(x = Visit,
                y = fit)
) +
  facet_grid(
    cols = vars(SNP),
    rows = vars(Period),
    labeller = labeller(
      SNP = snp.labs,
      Period = period.labs
    )
  ) +
  geom_line(aes(color=id)) +
  geom_point(aes(color=id)) +
  theme_bw() +
  labs(
    x = "Visit",
    y = "Probability of Severe Disease",
    title = "Subject-Specific Fitted Probability of Severe Disease"
  ) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none")
```

```{r glmm_viz_marginal_probability}
data.fitted.marginal <- data.frame(
  id = 1,
  Visit = 0:7,
  Period = rep(c(rep("A", 8), rep("B", 8)), 3),
  SNP = factor(0:2),
  scaled_age = mean(DF$scaled_age),
  sex = c(rep("male", 48), rep("female", 48))
)
probs <- predict(final.glmm, 
                 newdata = data.fitted.marginal, 
                 type_pred = "response", 
                 type = "marginal", 
                 re.form = NA, 
                 se.fit = TRUE)
data.fitted.marginal$fit <- probs$pred

snp.labs <- c("SNP=0", "SNP=1", "SNP=2")
names(snp.labs) <- c(0, 1, 2)

period.labs <- c("Period A", "Period B")
names(period.labs) <- c("A", "B")

ggplot(
  data = data.fitted.marginal,
  mapping = aes(x = Visit, y = fit, color=sex)
) +
  facet_grid(
    cols = vars(SNP),
    rows = vars(Period),
    labeller = labeller(
      SNP = snp.labs,
      Period = period.labs
    )
  ) +
  geom_line() +
  geom_point() +
    scale_color_manual(values=c("coral2", "dodgerblue4"),
                     labels=c("Female", "Male"),
                     name="Sex") +
  theme_bw() +
  labs(
    x = "Visit",
    y = "Probability of Severe Disease",
    title = "Marginal Probability of Severe Disease"
  ) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom")
```
