---
title: "EMLM Final Case Study"
author: "Colin Yip"
date: "2024-03-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(caret)
library(ggpubr)
library(lme4)
library(GLMMadaptive)
library(MASS)
```

#### Data loading
```{r}
load("Group_9.RData")
resp_var <- DF$Severity
```


#### Distribution of severity relative to covariates
Checking for distribution of severity/bias in severity between groups
```{r}
with(DF, tapply(Severity, list(Visit),
  mean,
  na.rm = TRUE
))
with(DF, tapply(Severity, list(Visit, sex),
  mean,
  na.rm = TRUE
))
with(DF, tapply(Severity, list(Visit, SNP),
  mean,
  na.rm = TRUE
))
with(DF, tapply(Severity, list(Visit),
  var,
  na.rm = TRUE
))
with(DF, tapply(Severity, list(Visit, sex),
  var,
  na.rm = TRUE
))
with(DF, tapply(Severity, list(Visit, SNP),
  var,
  na.rm = TRUE
))

bar_width <- 0.7

sex_labs <- c("Male", "Female")
names(sex_labs) <- c("male", "female")

sex_sev_dist <- ggplot(data = DF, mapping = aes(
  x = as.factor(Severity),
  fill = as.factor(Severity)
)) +
  scale_fill_manual(values = c("seagreen", "purple")) +
  facet_grid(
    cols = vars(sex),
    labeller = labeller(sex = sex_labs)
  ) +
  geom_bar(
    stat = "count",
    color = "black",
    width = bar_width
  ) +
  ggplot2::theme_bw() +
  labs(
    x = "",
    y = "Count",
    title = "Sex"
  ) +
  theme(
    legend.position = "None",
    plot.title = element_text(hjust = 0.5)
  ) +
  ylim(0, 1300)

period_sev_dist <- ggplot(data = DF, mapping = aes(
  x = as.factor(Severity),
  fill = as.factor(Severity)
)) +
  scale_fill_manual(values = c("seagreen", "purple")) +
  facet_grid(cols = vars(Period)) +
  geom_bar(
    stat = "count",
    color = "black",
    width = bar_width
  ) +
  ggplot2::theme_bw() +
  labs(
    x = "",
    y = "Count",
    title = "Period"
  ) +
  theme(
    legend.position = "None",
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ) +
  ylim(0, 1300)

snp_sev_dist <- ggplot(data = DF, mapping = aes(
  x = as.factor(Severity),
  fill = as.factor(Severity)
)) +
  scale_fill_manual(values = c("seagreen", "purple")) +
  facet_grid(cols = vars(SNP)) +
  geom_bar(
    stat = "count",
    color = "black",
    width = bar_width
  ) +
  ggplot2::theme_bw() +
  labs(
    x = "Severity",
    y = "Count",
    title = "SNP"
  ) +
  theme(
    legend.position = "None",
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ) +
  ylim(0, 1300)

joined_fig <- ggarrange(sex_sev_dist, snp_sev_dist, period_sev_dist,
  ncol = 3
)

annotate_figure(joined_fig,
  top = text_grob("Distribution of Severity Level (0, 1) by Categorical Predictor",
    face = "bold",
    size = 13
  )
)
```

Density of relative Severity factor representation as a function of age bins
```{r}
ggplot(
  data = DF,
  mapping = aes(x = Age, fill = as.factor(Severity))
) +
  facet_grid() +
  scale_fill_manual(values = c("seagreen", "purple")) +
  geom_density(alpha = 0.4) +
  ggplot2::theme_bw() +
  guides(fill = guide_legend(title = "Severity")) +
  labs(
    title = "Age Density by Severity Level",
    x = "Age",
    y = "Density"
  )
```
#### Mean structure of severity over visits
```{r}
# Means over time points
plot_ylim <- ylim(0, 0.55)

sex_grouped_data <- DF |>
  group_by(sex, Visit) |>
  summarize(Severity = mean(Severity))

sex_mean_structure <- ggplot(
  data = sex_grouped_data,
  mapping = aes(
    x = factor(Visit),
    y = Severity,
    group = sex,
    color = sex
  )
) +
  geom_line(stat = "summary", fun = "mean") +
  scale_color_manual(
    values = c("#F8766D", "#00BFC4"),
    labels = c("Male", "Female"),
    guide = guide_legend(direction = "horizontal")
  ) +
  ggplot2::theme_bw() +
  labs(
    title = "Sex",
    x = "",
    y = "Severity"
  ) +
  guides(fill = guide_legend(ncol = 2)) +
  theme(
    legend.position = c(0.6, 1.045),
    legend.margin = margin(c(1, 5, 1, 5)),
    legend.title = element_blank()
  ) +
  plot_ylim

snp_mean_structure <- ggplot(
  data = DF |>
    group_by(SNP, Visit) |>
    summarize(Severity = mean(Severity)),
  mapping = aes(
    x = factor(Visit),
    y = Severity,
    group = factor(SNP),
    color = factor(SNP)
  )
) +
  geom_line(stat = "summary", fun = "mean") +
  scale_color_manual(
    values = c("#f72585", "#7209b7", "#207bff"),
    guide = guide_legend(direction = "horizontal")
  ) +
  ggplot2::theme_bw() +
  labs(
    title = "SNP",
    x = "Visit",
    y = "Severity"
  ) +
  guides(fill = guide_legend(ncol = 3)) +
  theme(
    legend.position = c(0.65, 1.045),
    legend.margin = margin(c(1, 5, 1, 5)),
    legend.title = element_blank(),
    axis.title.y = element_blank()
  ) +
  plot_ylim

period_mean_structure <- ggplot(
  data = DF |>
    group_by(Period, Visit) |>
    summarize(Severity = mean(Severity)),
  mapping = aes(
    x = factor(Visit),
    y = Severity,
    group = Period,
    color = Period
  )
) +
  geom_line(stat = "summary", fun = "mean") +
  scale_color_manual(
    values = c("#e4572e", "#ffc914"),
    guide = guide_legend(direction = "horizontal")
  ) +
  ggplot2::theme_bw() +
  labs(
    title = "Period",
    x = "",
    y = "Severity"
  ) +
  theme(
    legend.position = c(0.75, 1.045),
    legend.margin = margin(c(1, 5, 1, 5)),
    legend.title = element_blank(),
    axis.title.y = element_blank()
  ) +
  plot_ylim

joined_mean_structure_fig <- ggarrange(sex_mean_structure, snp_mean_structure, period_mean_structure,
  ncol = 3
)

annotate_figure(joined_mean_structure_fig,
  top = text_grob("Proportion of Severe = 1 by Visit and Predictor",
    face = "bold",
    size = 13
  )
)
```

### Logistic mixed effects model
```{r}
# Change types of variables to factors
DF$Severity <- as.factor(DF$Severity)
DF$Period <- as.factor(DF$Period)
DF$sex <- as.factor(DF$sex)
DF$SNP <- as.factor(DF$SNP)

# Centering and scaling age, so coefficient interpretation becomes
# what is the change in the response relative to a 1SD change in the predictor
DF$scaled_age <- scale(DF$Age)
```

For `glmer`, Laplace approximation is not usable as it does not behave well with binary data. As such, we will use adaptive Gaussian quadrature where possible instead to approximate the integral $\int f(z)\phi(z)dz\approx\sum_{q=1}^{15}w_qf(z_q), \phi(z)\text{ being the multivariate normal},$ which is the likelihood contribution of each subject.

Approach assumes independent errors.

#### Setting up elaborate mean model and picking a random effects structure
```{r}
# Opting for mixed_model directly, as convergence more likely due to
# aGQ being more flexible to either high dimensional, or broadly more uneven/
# extreme likelihood function surfaces --> glmer ran into high eigenvalue ratios
# for the Hessian for increased interaction terms and/or increasing the number of
# random effects. glmer also ran into gradient errors unless a high
# (~1e5) number of iterations was run --> not very computationally efficient

# Check to see if glmer converges
model.0 <- glmer(
  Severity ~ Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age +
    (1 | id),
  data = DF,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa"),
  nAGQ = 15
)

# A penalized quasi-likelihood model runs into similar issues
model.pql <- glmmPQL(
  Severity ~ Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age,
  random = ~ 1 | id,
  family = binomial,
  data = DF
)

# Skipping checks with Laplace due to computational issues with binary variables

# Linear coefficients, random intercept, maximal
model.1 <- mixed_model(
  Severity ~ Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age,
  random = ~ 1 | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

# Linear coefficients, random intercept, random slope, maximal
model.2 <- mixed_model(
  Severity ~ Visit * Period +
    Visit * sex +
    Visit * SNP +
    Visit * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

# Quadratic visit, random intercept only, maximal
model.3 <- mixed_model(
  Severity ~ (Visit + I(Visit^2)) * Period +
    (Visit + I(Visit^2)) * sex +
    (Visit + I(Visit^2)) * SNP +
    (Visit + I(Visit^2)) * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age,
  random = ~ 1 | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

# Quadratic visit, random intercept and random slope, maximal
model.4 <- mixed_model(
  Severity ~ (Visit + I(Visit^2)) * Period +
    (Visit + I(Visit^2)) * sex +
    (Visit + I(Visit^2)) * SNP +
    (Visit + I(Visit^2)) * scaled_age +
    Period * sex +
    Period * SNP +
    Period * scaled_age +
    sex * SNP +
    sex * scaled_age +
    SNP * scaled_age,
  random = ~ Visit | id,
  data = DF,
  family = "binomial",
  control = list(nAGQ = 15)
)

if (all(
  model.1$converged, model.2$converged,
  model.3$converged, model.4$converged
)) {
  print("Achieved solution convergence in all models")
} else {
  print("One of the above models has not converged")
}

# Linear vs. quadratic to determine the correct fixed effects
anova(model.1, model.3)
# H_0: Coefficients for quadratic terms are 0, H_A being that they are != 0
# p-value >= 0.05, so we do not include quadratic terms and select from the linear models

# Linear comp (rand int vs rand int + rand slope)
random_effects_lrt <- anova(model.1, model.2)$LRT
0.5 * pchisq(random_effects_lrt, df = 1, lower.tail = FALSE) +
  0.5 * pchisq(random_effects_lrt, df = 2, lower.tail = FALSE)

# Between the linear models, rand int+rand slope (model.2) is signif. better
# Reject H_0: D=[[\sigma^2_int, 0][0,0]]
# Take H_A: D=[[\sigma^2_int, \sigma_{int,time}][\sigma_{int,time},\sigma^2_time]]

# Move forward with model.2
summary(model.2)
```

#### Simplification of elaborate mean model
```{r}
model.2.coefs <- summary(model.2)$coef_table
signif.coefs <- model.2.coefs[model.2.coefs[, 'p-value'] < 0.05, ]

# TODO: Write little blurb about complications of stepwise for mixed effects models



```
